---
title: "Coursera hopkins statistical inference"
author: "TD"
date: "02 May 2019"
---
# Settings
Setting up the enviroment
```{r setting up the environment, echo=TRUE}
rm(list=ls())
library(ggplot2)
library(dplyr)
```

# Data loading
Loading the data from ToothGrowth and inserting it into a tibble. In the tibble a new category (combination of supp and dose) is added.
```{r loading data, echo=TRUE}
df_tooth_growth <-  ToothGrowth

tbl_tooth_growth <- as_tibble(df_tooth_growth)
rm(df_tooth_growth)
tbl_tooth_growth <- mutate(tbl_tooth_growth, supplement_dose = paste(supp, dose, sep=', '))
```

# Data analysis

Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions.
First the data is summarized showing that the tibble is of size 60 x 4. It contains four data categories (length, supp, dose and supplement_dose). The len has a min of 4.2 and a max of 33.9. The dose has a min of 0.5, a max of 2 and a median of 1.

```{r summarizing data, echo=TRUE}
dim(tbl_tooth_growth)
head(tbl_tooth_growth)
summary(tbl_tooth_growth)

```
The data appearantly has growth length for teeth with  growth supplement (vitamine C) of two suppliers and three doses. The growth rate differences per dose are shown before.
```{r dose impact, echo=TRUE}
boxplot(tbl_tooth_growth$len ~ tbl_tooth_growth$supplement_dose)
```

And below it is seen what influence the doses from different suppliers have on the teeth growth.
```{r dose impact per supplier, echo=TRUE}
ggplot(data=tbl_tooth_growth, aes(x=dose, y=len)) +
    geom_point(aes(color=supp))
```

Then the hypothesis is tested that OJ is better than VC for a dose of 0.5. The p value is 0.003 which is lower than a 5% confidence interval and we therefore concluded that OJ is indeed better than VC (e.g. stronger teeth growth) for a dose of 0.5
```{r test hypothesis that OJ is better than VC for dose 0.5, echo=TRUE}
tbl_oj_5 <- filter(tbl_tooth_growth, dose==0.5, supp=='OJ')
tbl_vc_5 <- filter(tbl_tooth_growth, dose==0.5, supp=='VC')
t.test(tbl_oj_5$len, tbl_vc_5$len, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.05)$p.value
```

Next the hypothesis is tested that OJ is better than VC for a dose of 1.0. The p value is 0.0005 which is lower than a 5% confidence interval and we therefore concluded that OJ is indeed better than VC (e.g. stronger teeth growth) for a dose of 1.0
```{r test hypothesis that OJ is better than VC for dose 1.0, echo=TRUE}
tbl_oj_10 <- filter(tbl_tooth_growth, dose==1.0, supp=='OJ')
tbl_vc_10 <- filter(tbl_tooth_growth, dose==1.0, supp=='VC')
t.test(tbl_oj_10$len, tbl_vc_10$len, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.05)$p.value
```
Finally the hypothesis is tested that OJ is better than VC for a dose of 2.0. The p value is 0.5 which is higher than a 5% confidence interval and we therefore fail to reject the 0-hypothesis (e.g. VC is better than OJ) for a dose of 1.0
```{r test hypothesis that OJ is better than VC for dose 2.0, echo=TRUE}
tbl_oj_20 <- filter(tbl_tooth_growth, dose==2.0, supp=='OJ')
tbl_vc_20 <- filter(tbl_tooth_growth, dose==2.0, supp=='VC')
t.test(tbl_oj_20$len, tbl_vc_20$len, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.05)$p.value
```
Now we compare teeth growth for a dose of 0.5 with a dose of 1.0 and a dose of 1.0 with a dose of 2.0.
It is seen that the confidence intervals for the hypothesis that a dose of 1.0 is better than a dose of 0.5 which has a p-value of 6e-8 and is therefore confirmed. The hypothesis that a dose of 2.0 is better than a dose of 1.0 has a p-value of 1e-5 and is therefore confirmed.
```{r higher dose is better than lower dose}
tbl_5 <- filter(tbl_tooth_growth, dose==0.5)
tbl_10 <- filter(tbl_tooth_growth, dose==1.0)
tbl_20 <- filter(tbl_tooth_growth, dose==2.0)
t.test(tbl_10$len, tbl_5$len, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.05)$p.value
t.test(tbl_20$len, tbl_10$len, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.05)$p.value
```